---
title: "rrtools:"
subtitle: "Adapting R package development tools for Reproducible Research"
author: "Anna Krystalli - @ Sheffield R Users Group"
date: "2018/04/10"
output:
  xaringan::moon_reader:
    css: ["default", "assets/css/little_ak.css", "assets/css/little_ak-fonts.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, echo=F, message=F, warning=F}
library(magick)
library(dplyr)
```


# Why bother?

## - Convention
## - Tools


---

# Convention

> It’s like agreeing that we will all drive on the left or the right. A hallmark of civilization is following conventions that constrain your behavior a little, in the name of public safety.

**Jenny Bryan** on [Project-oriented workflows](https://www.tidyverse.org/articles/2017/12/workflow-vs-script/)

---

## Hierarchy of convention

- **Basic project >** Carpentries [Project Management With RStudio](https://swcarpentry.github.io/r-novice-gapminder/02-project-intro/) 
- **Package >**: _[R packages](http://r-pkgs.had.co.nz/intro.html) by Hadley Wickham_ 
- **Reproducible Research compendium >**: [Rrtools](https://github.com/benmarwick/rrtools)


---

# Tools

### Leverage tools and functionality for **R package development**
- manage dependencies
- make functionality available
- document functionality
- validate functionality
- version contol your project


[`devtools`](https://www.tidyverse.org/articles/2017/11/usethis-1.0.0/), [`usethis`](https://www.tidyverse.org/articles/2017/11/usethis-1.0.0/), Rstudio

---

## A reproducible research compendium:

[_Wickham, H. (2017) Research compendia. Note prepared for the 2017 rOpenSci Unconf_](https://docs.google.com/document/d/1LzZKS44y4OEJa4Azg5reGToNAZL0e0HSUwxamNY7E-Y/edit#heading=h.blggi16hdosp)

> ### Goals

> - Provide workflow tools that make life easier for analysts. **Only a small upfront investment should be required to get a significant payoff**, and the tools should **grow** with you as your **analyses get more complex**.

> - Make it **easier to transition from ephemeral low-stakes to high-stakes functions** at the core of an analysis. In particular, it should be **easy to document and test important functions**.


---

## A reproducible research compendium:

[_Wickham, H. (2017) Research compendia. Note prepared for the 2017 rOpenSci Unconf_](https://docs.google.com/document/d/1LzZKS44y4OEJa4Azg5reGToNAZL0e0HSUwxamNY7E-Y/edit#heading=h.blggi16hdosp)

> ### Goals


> - **Ensure that an analysis project is a package, without requiring any knowledge of package development** (so that newcomers are not overwhelmed with too many new things to learn), while laying the foundation for projects based solely around code reuse.

> - Should be **highly opinionated about directory names and conventions**; should be not opinionated about the packages you use for the analysis. 


---

### Proposal


#### A ***minimal*** analysis project would consist of:

1. An **`scripts/`** directory that contains R scripts (`.R`), notebooks (`.Rmd`), and intermediate data.

2. A **`DESCRIPTION`** file that provides metadata about the compendium. Most importantly, it would list the packages needed to run the analysis. Would contain field to indicate that this is an analysis, not a package.


--- 


#### A ***reproducible*** analysis project would also contain:

1. An **`R/`** directory which contains R files that provide high-stakes functions.

1. A **`data/`** directory which contains high-stakes data.

1. A **`tests/`** directory that contains unit tests for the code and data.

1. A **`vignettes/`** directory that contains high-stakes reports.

1. A **build system** which executes the contents of `scripts/` and `vignettes/`, producing final data and reports.

---

#### Some other components would be automatically added/generated:

1. A **`.Rbuildignore`** file that ignores the `scripts/` directory.  

1. A **`man/`** directory which contains 88roxygen2-generated documentation** for the reusable functions and data.

#### A ***shareable*** reproducible analysis project would also:

- Use Git + GitHub (or other public Git host)

- Use Travis or other continuous integration service

- Capture the computational environment so it can easily be recreated on a different computer. This involves at a minimum capturing package versions, but might also include capturing R version, and other external dependencies.

---

## `rrtools`

> The goal is to provide a continuum of reproducibility so the newcomer can start off easy and graduate to more reproducible, but more complex, workflows over time.


### install `rrtools`

```{r, eval=FALSE}
devtools::install_github("benmarwick/rrtools")
```

---

### Create new compendium

```{r, eval=FALSE}
rrtools::use_compendium("../../dummy/rrtoolsdemo")
```


```{r, echo=FALSE}
magick::image_read("assets/use_compendium.png")
```

---

### Resulting compendium

```{r, echo=FALSE}
magick::image_read("assets/compendium1.png")
```

---

## `DESCRIPTION`

### Add basic metadata

- Package title
- Package description
- Author names

etc.

---

### Manage dependencies


#### Add CRAN dependencies

```{r, eval=FALSE}
usethis::use_package("ggplot2")
usethis::use_package("dplyr")
```


#### Add development version of a dependency (ie from GitHub)
```{r, eval=FALSE}
use_dev_package("ggplot2")
```

#### Add a suggested package 
(packages used in testing or in vignettes but required to run compendium functions)
```{r, eval=FALSE}
usethis::use_package("gapminder", type = "Suggests")
```

---

## `R/` directory 

Keep all scripts containing core functions in the `/R` directory. These can be made available by installing and loading as any other package


```{r}
usethis::use_r("data-processing")
```

```{r, echo=FALSE}
magick::image_read("assets/data-processing-fun.png") %>% image_resize("600x700")
```


---

### Document functions using [Roxygen2](http://kbroman.org/pkg_primer/pages/docs.html) notation

This allows documentation to be generated automatically on build

```{r, echo=FALSE}
magick::image_read("assets/function_help.png") %>% image_resize("600x700")
```

---




use `<pkname>::<function-name>` to specify functions imported from other packages namespaces. This helps:

- accurate specification of function to be used regardless of package loading sequence.
- ability to easily distinquish between imported functions and your own functions in your code.


---

## Version control

Most commonly through `git`

> START WITH: [Happy Git and GitHub for the useR](http://happygitwithr.com/)


```{r, eval=FALSE}
usethis::use_git()
```


---

## Collaboration through github

### Create repository


```{r, eval=FALSE}
rrtools::use_github()
```


```{r, echo=FALSE}
magick::image_read("assets/github.png") %>% image_resize("300x700")
```

```{r, echo=FALSE}
magick::image_read("assets/github_DES.png") %>% image_resize("400x700")
```

---

You'll need a Personal Authorisation Token (PAT).

```{r}
usethis::browse_github_pat()
```

will open up the GitHub panel to generate your PAT. 

---

Copy it and paste it into your `.Renviron` file as system variable `GITHUB_PAT`.

```{r, eval=F}
usethis::edit_r_environ()
```

Use `edit_r_environ()` to open and edit your `.Renviron` file


```{r, echo=FALSE}
magick::image_read("assets/GITHUB_PAT.png")
```


---

#### README

READMEs are the landing page of any repository on GitHub. 

The `rrtools` version of `use_readme_md()` creates a README template appropriate for a repository associated with a Research Project.

Render each time you make changes to `README.Rmd` so `README.md` is updated.

```{r, eval=FALSE}
rrtools::use_readme_rmd()
```

Also creates:

- `CONTRIBUTING.md`: Edit with guidance on contribution
- `CONDUCT.md`: Code of conduct associated with the project

---
```{r, echo=FALSE}
magick::image_read("assets/README.png") %>% image_resize("400x700")
```


---

### License

```{r, eval=FALSE}
usethis::use_mit_license("Anna Krystalli")
```


```{r, echo=FALSE}
magick::image_read("assets/LICENSE.png") %>% image_resize("200x700")
```


```{r, echo=FALSE}
magick::image_read("assets/LICENSEmd.png") %>% image_resize("400x700")
```

---

## Maintaining code integrity

---

### Testing


If you add functions in `R/`, include tests to ensure they function as intended

```{r, eval=FALSE}
usethis::use_testthat()
```

```{r, echo=FALSE}
magick::image_read("assets/add_testthat.png") %>% image_resize("400x700")
```



---

Create tests.R in `tests/testhat/` and check <http://r-pkgs.had.co.nz/tests.html> for template

```{r, eval=FALSE}
use_test("data-processing")
```

```{r, echo=FALSE}
magick::image_read("assets/add_test.png") %>% image_resize("400x700")
```

---

### Continuous Integration using Travis

```{r, eval=FALSE}
rrtools::use_travis(docker = F)
```

Once your tests are set up, you can ensure that changes to the code base do not break functionality by using continuous integration.

See Julia Silge's [A BEGINNER'S GUIDE TO TRAVIS-CI FOR R](https://juliasilge.com/blog/beginners-guide-to-travis/) for a gentle intro.

---

## Controlling environment with containers


```{r, eval=FALSE}
rrtools::use_dockerfile()
```


-   Creates a basic Dockerfile using [`rocker/verse`](https://github.com/rocker-org/rocker) as the base image
    - R
    - the [tidyverse](http://tidyverse.org/)
    - RStudio
    - pandoc
    - LaTeX
    so compendium build times are very fast on travis
-   version of R in your rocker container will match the version used when you run this function (e.g., `rocker/verse:3.4.0`)


---

## Vignettes

> Long form documentation

### Analyses
### Setup instructions
### Functionality demos
### Reports

---

### bookdown for publications

More on this from Dan.


--- 

### Rstudio for building and testing your code

```{r, echo=FALSE}
magick::image_read("assets/rstudio_build.png") %>% image_resize("400x700")
```
